Modelling the City of Melbourne Pedestrian Data

### 2017-05-07 presentation/proposal_pres.Rmd
more slides - still WIP

```{r, echo = FALSE, fig.height=5, fig.retina = 2}
ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude), data = ped_loc, colour = "red", alpha = 0.7) +
              labs(title = "Melbourne Pedestrian Sensor Locations") +
              xlab(NULL) + ylab(NULL) +
              theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5),
                    axis.text.x = element_blank(), axis.text.y = element_blank(),
                    axis.ticks = element_blank())
```

@@

The City of Melbourne has `r nrow(ped_loc)` sensors installed around Melbourne's CBD  
- These sensors are positioned under an awning or street pole, and count the number of people passing each hour

@@

  and many other various ways..

# What does the data look like?

```{r}
glimpse(ped_df)
```

```{r, echo = FALSE}
      ped_df$Year <- as.factor(ped_df$Year)
      ped_df$Month <- as.factor(ped_df$Month)
      ped_df$Month <- factor(ped_df$Month, levels(ped_df$Month)[c(5,4,8,1,9,7,6,2,12,11,10,3)])
      ped_df$Time <- as.factor(ped_df$Time)
      ped_df$Timeplot <- as.integer(ped_df$Time)
      ped_df$MTime <- as.integer(ped_df$Time) + ((as.integer(ped_df$Mdate)-1) * 24)
      ped_df$Day <- as.factor(ped_df$Day)
      ped_df$Day <- factor(ped_df$Day, levels(ped_df$Day)[c(2,6,7,5,1,3,4)])

@@

@@

@@

>>>>

```{r, echo = FALSE, fig.height=5, fig.retina = 3}
ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude), data = ped_loc, colour = "red", alpha = 0.7) +
              labs(title = "Melbourne Pedestrian Sensor Locations") +
              xlab(NULL) + ylab(NULL) +
              theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5),
                    axis.text.x = element_blank(), axis.text.y = element_blank(),
                    axis.ticks = element_blank())
```

@@

The City of Melbourne has `r nrow(ped_loc)` sensors installed around Melbourne's CBD  
- These sensors are positioned under an awning or street pole, and count the number of people passing each hour
- The data is available as a `.csv` at https://data.melbourne.vic.gov.au/, and is updated monthly

@@

  and many other various ways..
---

@@

```{r, echo = FALSE}
      ped_df$Year <- as.factor(ped_df$Year)
      ped_df$Month <- as.factor(ped_df$Month)
      ped_df$Month <- factor(ped_df$Month, levels(ped_df$Month)[c(5,4,8,1,9,7,6,2,12,11,10,3)])
      ped_df$Time <- as.factor(ped_df$Time)
      ped_df$Timeplot <- as.integer(ped_df$Time)
      ped_df$MTime <- as.integer(ped_df$Time) + ((as.integer(ped_df$Mdate)-1) * 24)
      ped_df$Day <- as.factor(ped_df$Day)
      ped_df$Day <- factor(ped_df$Day, levels(ped_df$Day)[c(2,6,7,5,1,3,4)])
      for (i in 1:dim(ped_loc)[1])
      {
      ped_df$Lat[ped_df$Sensor_ID == ped_loc$`Sensor ID`[i]] <- ped_loc$Latitude[i]
      ped_df$Lon[ped_df$Sensor_ID == ped_loc$`Sensor ID`[i]] <- ped_loc$Longitude[i]
      }

@@

```{r, echo = FALSE}
head(ped_df[, -c(11,12)], 1)
```

@@

# Modelling

- Main purpose of the model is to forecast pedestrian counts
- Data used for estimatation (training dataset) will be the period between 2014-2016
  + The data before 2014 is not stable, and also lacks data from newly installed sensors
- Data used to evaluate the forecasting performance will be the 2017 data available

- Due to the multiple seasonal patterns in the time series, typical models like AR, ARMA and ARIMA do not work very well
  + Modelling using other sources of data is being explored, in particular weather data
  + Alternative models, such as a Multiple Seasonal (MS) process available through the `forecast` R package, should also be investigated

## Visualisation: Currently available
#### Official City of Melbourne data visualisation

```{r, echo = FALSE, out.height= "400px", out.width = "540px"}
include_graphics("../img/ped_vis_melb.PNG")
```

---
class: center, middle

## Visualisation

With this project, we want to improve the way we visualise Melbourne's pedestrian activity, and use a predicitive model to get a better understanding of how Melbourne operates

```{r, echo = FALSE}
ggmap(melb) + geom_point(aes(x = Lon, y = Lat, size = Hourly_Counts), alpha = 0.02, data = filter(ped_df, Year == "2016", Month == "March", Time == "9"), colour = "violetred") + scale_size_area(max_size = 10) + guides(colour = guide_legend(override.aes = list(alpha = 1)))
```
